Scalar and Representative Observables for Polarimetric SAR Data

### 2013-07-24 paper_dev/article_pp.tex
updated for Ian changes

%opening
\title{
  Scalar and Representative Models for Polarimetric SAR Data
}

@@

\markboth{Transactions on Geoscience \& Remote Sensing,~Vol.~?, No.~?, ?~2013}%
{ Le \MakeLowercase{\textit{et al.}}:  Scalar and Representative Models for Polarimetric SAR Data}

@@

%ifdef FINAL
\begin{figure}[h]
\centering
\begin{tabular}{c}
	\subfloat[multi-look SAR dispersion]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_multi_look_sar_dispersion_pdf.eps} 	
		 \label{multi_look_dispersion}
	} 
	\hfill	
	\subfloat[multi-look SAR contrast]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_multi_look_sar_contrast_pdf.eps} 	
		 \label{multi_look_contrast}
	}
\end{tabular}
\caption{Multi-Look SAR dispersion and contrast: modelled response matches very well with real-life captured data.}
\label{fig:verify_multi_look_SAR_dispersion_contrast_models}
\end{figure}
%endif
%ifdef REVIEW
\begin{figure}[h!]
\centering
\begin{tabular}{c}
	\subfloat[multi-look SAR log-distance]{
		 \epsfxsize=3in
		 \epsfysize=3in
		 \epsffile{images/verify_multi_look_sar_dispersion_pdf.eps} 	
		 \label{multi_look_dispersion}
	} 
	\hfill	
	\subfloat[multi-look SAR contrast]{
		 \epsfxsize=3in
		 \epsfysize=3in
		 \epsffile{images/verify_multi_look_sar_contrast_pdf.eps} 	
		 \label{multi_look_contrast}
	}
\end{tabular}
\caption{Multi-Look SAR dispersion and contrast: modelled response matches very well with real-life captured data.}
\label{fig:verify_multi_look_SAR_dispersion_contrast_models}
\end{figure}
%endif

@@

%ifdef FINAL
\begin{figure}[h]
\centering
\begin{tabular}{c}
	\subfloat[Handling over-sampling practice in AIRSAR part-pol dataset]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/handling_airsar_oversampling_practice_part_pol_log_distance.eps} 	
		 \label{sar}
	} 
	\hfill	
	\subfloat[Handling over-sampling practice in AIRSAR full-pol dataset]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/handling_airsar_oversampling_practice_full_pol_log_distance.eps} 	
		 \label{part_pol}
	}   
\end{tabular}
\caption{AIRSAR Flevoland also exhibits phenomena of over-sampling practice, through to a lesser extent than the RADARSAT2 data.}
\label{fig:handling_airsar_oversampling_practice_full_pol}
\end{figure}
%endif
%ifdef REVIEW
\begin{figure}[h!]
\centering
\begin{tabular}{c}
	\subfloat[Handling over-sampling practice in AIRSAR part-pol dataset ]{
		 \epsfxsize=3in
		 \epsfysize=3in
		 \epsffile{images/handling_airsar_oversampling_practice_part_pol_log_distance.eps} 	
		 \label{sar}
	} 
	\hfill	
	\subfloat[Handling over-sampling practice in AIRSAR full-pol dataset]{
		 \epsfxsize=3in
		 \epsfysize=3in
		 \epsffile{images/handling_airsar_oversampling_practice_full_pol_log_distance.eps} 	
		 \label{part_pol}
	}   
\end{tabular}
\caption{AIRSAR Flevoland also exhibits phenomena of over-sampling practice, through at a lesser extend than the RADARSAT 2 data.}
\label{fig:handling_airsar_oversampling_practice_full_pol}
\end{figure}
%endif

@@

Let us begin by noting a few theoretical properties of the proposed statistical model.
First, the use of covariance matrix log-determinant may be related to the standard eigen-decomposition method of the POLSAR covariance matrices.
In fact, the log-determinant can also be computed as the sum of log-eigenvalues.
Specifically $\ln{|M|} = \sum \ln{\lambda_M}$ where $\lambda_M$ denotes all the eigenvalues of M.
Thus similar to other eigenvalue based approach (e.g. entropy/anisotropy, ...),
  the models presented here are invariant to polarization basis transformations.

Second, the model is developed for the POLSAR covariance matrix.
However, since the POLSAR coherency matrix is related to the covariance matrix via a unitary transformation which preserves the determinant,
  the model should also be applicable to the coherency matrix.

The model is far from complete.
 It reduces the multi-dimensional POLSAR data to a scalar value.
While this is probably desirable for a wide range of application,
  such a reduction is unlikely to be lossless.
Thus %to better understand POLSAR data
the use of this technique could be complemented with some high-dimensional POLSAR target-decomposition techniques (e.g. the Freeman Durden decomposition \cite{Freeman_1998_TGRS_963} or the entropy/anisotropy decomposition \cite{Cloude_1997_TGRS_68} or similar).

However the proposed model is promising.
Even though initially developed for partial and monostatic POLSAR data,
  it was shown to be applicable to traditional SAR data as well.
Since the models assumptions are quite minimal, they may also be found to apply to bi-static and interferometric data, although that would require significant further investigation.

The theoretical models may also provide an alternative derivation for the widely used likelihood test statistics in POLSAR.
In view of the models given in Eqns \ref{eqn:determinant_distribution} \& \ref{eqn:log_determinant_distribution},
  the likelihood test statistics exposed in \cite{Conradsen_2003_TGRS_4} and rewritten in Eqns \ref{eqn:ori_likelyhood_stats} \& \ref{eqn:log_likelyhood_stats}
can be simulated as:
\begin{align*}
  \ln{Q} &\sim  k + L_x \Lambda^d_{L_x} + L_y \Lambda^d_{L_y} - (L_x + L_y) \Lambda^d_{(L_x + L_y)} \\
  Q &\sim e^k \frac{(\chi^d_{L_x})^{L_x} \cdot (\chi^d_{L_y})^{L_y}}{(\chi^d_{L_x + L_y})^{L_x + L_y}}   
\end{align*}
where $k = d \left[ (L_x + L_y) \ln(L_x + L_y) - L_x \ln{L_x} - L_y \ln{L_y} \right]$.
As a by-product of this exact derivation, the current paper also proposed several simpler discrimination measures for the common case of $L_x=L_y$.

Similar to the way that other measures of distance can be used to derive POLSAR classifiers \cite{Lee_1999_TGRS}, change detectors \cite{Conradsen_2003_TGRS_4}, edge detectors \cite{Schou_2003_TGRS_20} or other clustering and speckle filtering techniques \cite{Le_2010_ACRS} \cite{Le_2011_ACRS}, 
new detection, classification, clustering or speckle filtering algorithms can be derived using the models presented in this paper.
%***IVM remove this single sentence, it detracts from the argument...
%Moreover many of the existing techniques in SAR may be adaptable for POLSAR data, such as visual evaluation of speckle filters.

In evaluating SAR speckle filters, the intensity ratio is widely used.
Specifically, in its original multiplicative domain, the ratio of the filtered output and the noisy input image represents the noise being removed.
Assuming a perfect filtering condition, the ratio image should only contain random noise.
Thus a commonly used visual evaluation is to simply plot the ratio image and determine if any image features have also been removed.

@@

To briefly evaluate the performance of 3$\times$3 and 5$\times$5 boxcar POLSAR filters on the AIRSAR Flevoland partial polarimetric data (HH-HV),
  a square 700$\times$700 pixel patch is extracted from the AIRSAR dataset,
  and the two POLSAR speckle filters are then applied to the patch.
\footnote{While the boxcar filters may not stay in the frontier of POLSAR speckle filtering,
  they are workable here since the focus of this experiment is not on finding the best POLSAR speckle filters available.
  Rather its purpose is to show a quick example of how an existing technique in SAR data processing can be easily adapted for POLSAR data.} 
The log-determinant images of the filtered outputs are displayed in Fig. \ref{fig:visual_eval_part_pol_boxcar_speckle_filters_3x3_vs_5x5}.
At the same time, the residual is computed for both filters, and the images are also displayed in the same figure.

@@

Ian version: There are many similar SAR data processing techniques such as  the usage of  the mean and variance of the ratio image to quantitatively evaluate speckle filters.
Evidently these techniques can also be adapted for POLSAR.
Other examples of such techniques include existing SAR speckle filters, target detectors, land-cover classifiers and edge detectors.

Hai version: (Ian have not seen this yet) There are many similar SAR data processing technique that may be adaptable to POLSAR.
Examples in the same topic of evaluating speckle filters include
  the use of the mean and variance of the ratio image, or ENL evaluation to quantitatively evaluate POLSAR speckle filters.
Other topics that may also be connected include: speckle filtering, target detection, surface cover classification, edge detection ... and so on.

In fact, at the time of this writing, the field of SAR is much more developed than the field of POLSAR.
The two fields however are quite separated,
  as many techniques applicable to SAR are not directly applicable to POLSAR.
This work may build the bridge between these two separated fields,
  inviting potential extension of various SAR processing techniques on POLSAR data.
Here a specific example is shown
  where, similar to the way ratio was commonly used to carry out visually evaluation of SAR speckle filters,
  the newly proposed determinant ratio and its log-transformed version (i.e. contrast) are shown capable of visually evaluating POLSAR speckle filters.
In general, we believe there are many more unexplored possibilities for such applications
  which is only limited by our imagination.

@@

A number of scalar statistical models for the determinant of the POLSAR covariance matrix are proposed and validated in this paper,
  resulting in several consistent discrimination measures for POLSAR applications.
The theoretical models are shown to be powerful in that
  not only can they provide alternative and simple explanations to a range of theoretical concepts such as POLSAR test statistics or ENL estimation,
  but they can also be considered multi-dimensional extensions of traditional models for one-dimensional SAR data.
They are also practically useful since they lead to several consistent discrimination measures.
Moreover, compared to other scalar statistical models for POLSAR, the proposed models are excellent representations of the multi-dimensional POLSAR data.
Consequently, the derived discrimination measures may be employed in a wide range of applications where a scalar number is required to represent the complex multi-dimensional POLSAR data.
Thus the models are envisioned to form a bridge inviting the adaptation of many existing SAR data processing techniques for POLSAR data.

@@

In this appendix, the following results for SAR intensity $I$ are shown to be special cases of the results given in this paper for the determinant of the POLSAR covariance matrix $det|C_v|$.
Specifically, the following results extend from the authors previous work on single-look SAR \cite{Le_2013_TGRS_SAR_MSE}, i.e. $d=L=1$, which is considerd a special case. We can state the following:
\begin{eqnarray}
  I &\sim& \bar{I} \cdot pdf \left[ e^{-R} \right] \\
  \log_2{I} &\sim& \log_2{\bar{I}} + pdf \left[ 2^{D-2^D} \right] \\
  \frac{I}{\bar{I}} = \mathbb{R} &\sim& pdf \left[ e^{-R} \right]  \\
  \log_2{I} - \log_2{\bar{I}} = \mathbb{D} &\sim& pdf \left[ 2^De^{-2^D}\ln2 \right]\\
  \log_2{I_1} - \log_2{I_2} = \mathbb{C} &\sim& pdf \left[ \frac{2^c}{(1+2^c)^2} \ln2 \right] \\
  avg(\mathbb{D}) &=& -\gamma / \ln{2} \\
  var(\mathbb{D}) &=& \frac{\pi^2}{6} \frac{1}{ \ln^2{2}} \\
  mse(\mathbb{D}) &=& \frac{1}{\ln^2{2}}( \gamma^2 + \pi^2/6 ) = 4.1161 
\end{eqnarray}
but also the following well-known results are considered for multi-look SAR, i.e. $d=1,L>1$:
  \begin{eqnarray}
I &\sim& pdf \left[ \frac{L^L I^{L-1} e^{-LI/\bar{I}}}{\Gamma(L) \bar{I}^L} \right] \\
N = \ln{I} &\sim& pdf \left[ \frac{L^L}{\Gamma(L)} e^{L(N-\bar{N})-Le^{N-\bar{N}}} \right]
  \end{eqnarray}
It will be shown that all of these results are special cases of the result derived previously and rewritten below:
\begin{eqnarray}
  |C_v| &\sim& \frac{|\Sigma_v|}{(2L)^d} \prod^{d-1}_{i=0} \chi^2(2L-2i)  \label{eqn:polsar_det_cov_dist} \\
  \ln{|C_v|} &\sim& \ln{|\Sigma_v|} + \sum^{d-1}_{i=0} \Lambda(2L-2i) - d \cdot \ln{2L} \label{eqn:polsar_log_det_cov_dist} 
\end{eqnarray}
\begin{eqnarray}
  \frac{|C_v|}{|\Sigma_v|} = \mathbb{R} &\sim& \frac{1}{(2L)^d} \prod^{d-1}_{i=0} \chi^2(2L-2i) \label{eqn:polsar_ratio_det_cov_dist} \\
  \ln{|C_v|} - \ln{|\Sigma_v|} = \mathbb{D} &\sim& \sum^{d-1}_{i=0} \Lambda(2L-2i) - d \cdot \ln{2L} \label{eqn:polsar_dispersion_log_det_cov_dist} \\ 
  \ln{|C_{1v}|} - \ln{|C_{2v}|} = \mathbb{C} &\sim& \sum^{d-1}_{i=0} \Delta(2L-2i)
\end{eqnarray}
\begin{eqnarray}
  avg(\mathbb{D}) &=& \sum^{d-1}_{i=0} \psi^0(L-i) - d \cdot \ln{L} \label{eqn:polsar_dispersion_averages} \\
  var(\mathbb{D}) &=& \sum^{d-1}_{i=0} \psi^1(L-i) \label{eqn:polsar_dispersion_variance} \\
  mse(\mathbb{D}) &=& \left[ \sum^{d-1}_{i=0} \psi^0(L-i) - d \cdot \ln{L} \right]^2 +  \sum^{d-1}_{i=0} \psi^1(L-i) \label{eqn:polsar_dispersion_mse}
\end{eqnarray}

@@

The output of the first phase, in the general case, is applicable to multi-look SAR data, where $d=1$ but $L>1$.
The second phase simplifies the multi-look results into single-look results, which will match those presented in our previous work \cite{Le_2013_TGRS_SAR_MSE}.
Mathematically, it means setting $L=1$ in the multi-look result
  and converting from the natural logarithmic domain used in this paper to the base-2 logarithm used in \cite{Le_2013_TGRS_SAR_MSE} (base-2 was chosen in the previous paper to simplify the computation).

@@

Now setting $L=1$, these results become:
\begin{eqnarray}
    I &\sim& pdf \left[ \frac{ e^{x/\bar{I}}}{ \bar{I}} \right] \\
    \frac{I}{\bar{I}} = \mathbb{R} &\sim& pdf \left[ e^{-x} \right] 
\end{eqnarray}
which is the same as stated in \cite{Le_2013_TGRS_SAR_MSE}, demonstrating that the previous work is a special case of the more generic POLSAR forms.

@@

Setting $L=1$ for Single-Look SAR we have
\begin{equation}
  \mathbb{D} \sim pdf \left[ e^{x-\gamma-e^{x-\gamma}} \right]
\end{equation}
due to: $\psi^0(1)=-\gamma$ and $\Gamma(1)=1$
with $\gamma$ being the Euler Mascheroni constant (which equals $0.5772$). 
In base-2 logarithm domain, invoking the variable change theorem:
\begin{eqnarray*}
  \mathbb{D}_2 &=& \log_2{I} - avg(\log_2{I}) = \mathbb{D}/\ln{2} \\
  \mathbb{D}_2 &\sim& pdf \left[ e^{x-\gamma-e^{x-\gamma}} \cdot \frac{dx}{dt} \right]_{x=t \cdot \ln2}
\end{eqnarray*}
Thus we have
\begin{equation}
  \mathbb{D}_2 \sim pdf \left[ e^{-(2^xe^{-\gamma})} (2^xe^{-\gamma}) \ln2 \right]
\end{equation}
which is consistent with the results found in our previous work \cite{Le_2013_TGRS_SAR_MSE}.

>>>>

%opening
\title{
  Scalar and Representative Observables for Polarimetric SAR Data
}

@@

\markboth{IEEE Journal on Selected Topic in Applied Earth Observation and Remote Sensing,~Vol.~?, No.~?, ?~2013}%
{ Le \MakeLowercase{\textit{et al.}}:  Scalar and Representative Observables for Polarimetric SAR Data}

@@

%ifdef FINAL
\begin{figure}[h]
\centering
\begin{tabular}{c}
	\subfloat[multi-look SAR dispersion]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_multi_look_sar_dispersion_pdf.eps} 	
		 \label{multi_look_dispersion}
	} 
	\hfill	
	\subfloat[multi-look SAR contrast]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_multi_look_sar_contrast_pdf.eps} 	
		 \label{multi_look_contrast}
	}
\end{tabular}
\caption{Multi-Look SAR dispersion and contrast: modelled response matches very well with real-life captured data.
It should be noted that an even better match than what is shown here is presented in Fig. \ref{fig:handling_airsar_oversampling_practice_full_pol} where a better ENL estimation is employed.}
\label{fig:verify_multi_look_SAR_dispersion_contrast_models}
\end{figure}
%endif
%ifdef REVIEW
\begin{figure}[h!]
\centering
\begin{tabular}{c}
	\subfloat[multi-look SAR log-distance]{
		 \epsfxsize=3in
		 \epsfysize=3in
		 \epsffile{images/verify_multi_look_sar_dispersion_pdf.eps} 	
		 \label{multi_look_dispersion}
	} 
	\hfill	
	\subfloat[multi-look SAR contrast]{
		 \epsfxsize=3in
		 \epsfysize=3in
		 \epsffile{images/verify_multi_look_sar_contrast_pdf.eps} 	
		 \label{multi_look_contrast}
	}
\end{tabular}
\caption{Multi-Look SAR dispersion and contrast: modelled response matches very well with real-life captured data.
It should be noted that an even better match than what is shown here is presented in Fig. \ref{fig:handling_airsar_oversampling_practice_full_pol} where a better ENL estimation is employed.}
\label{fig:verify_multi_look_SAR_dispersion_contrast_models}
\end{figure}
%endif

@@

%ifdef FINAL
\begin{figure}[h]
\centering
\begin{tabular}{c}
	\subfloat[Handling over-sampling practice in AIRSAR part-pol dataset]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/handling_airsar_oversampling_practice_part_pol_log_distance.eps} 	
		 \label{sar}
	} 
	\hfill	
	\subfloat[Handling over-sampling practice in AIRSAR full-pol dataset]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/handling_airsar_oversampling_practice_full_pol_log_distance.eps} 	
		 \label{part_pol}
	}   
\end{tabular}
\caption{AIRSAR Flevoland also exhibits phenomena of over-sampling practice, through to a lesser extent than the RADARSAT2 data.}
\label{fig:handling_airsar_oversampling_practice_full_pol}
\end{figure}
%endif
%ifdef REVIEW
\begin{figure}[h!]
\centering
\begin{tabular}{c}
	\subfloat[Handling over-sampling practice in AIRSAR part-pol dataset ]{
		 \epsfxsize=3in
		 \epsfysize=3in
		 \epsffile{images/handling_airsar_oversampling_practice_part_pol_log_distance.eps} 	
		 \label{sar}
	} 
	\hfill	
	\subfloat[Handling over-sampling practice in AIRSAR full-pol dataset]{
		 \epsfxsize=3in
		 \epsfysize=3in
		 \epsffile{images/handling_airsar_oversampling_practice_full_pol_log_distance.eps} 	
		 \label{part_pol}
	}   
\end{tabular}
\caption{AIRSAR Flevoland also exhibits phenomena of over-sampling practice, through at a lesser extent than the RADARSAT 2 data.}
\label{fig:handling_airsar_oversampling_practice_full_pol}
\end{figure}
%endif

@@

Let us begin by noting a few theoretical properties of the proposed statistical models.
First, the use of covariance matrix log-determinant may be related to the standard eigen-decomposition method of the POLSAR covariance matrices.
In fact, the log-determinant can also be computed as the sum of log-eigenvalues.
Specifically $\ln{|M|} = \sum \ln{\lambda_M}$ where $\lambda_M$ denotes all the eigenvalues of M.
Thus similar to other eigenvalue based approach (e.g. entropy/anisotropy, ...),
  the models presented here are invariant to polarization basis transformations.

Second, the models are developed for the POLSAR covariance matrix.
However, since the POLSAR coherency matrix is related to the covariance matrix via a unitary transformation which preserves the determinant,
  the model should also be applicable to the coherency matrix.

It should be noted that the models are far from complete.
% It reduces the multi-dimensional POLSAR data to a scalar value.
%While this is probably desirable for a wide range of application,
%  such a reduction is unlikely to be lossless.
While it is desirable to reduce the multi-dimensonal POLSAR data to a scalar value for many applications,
  such a reduction is unlikely to be lossless.  
Thus %to better understand POLSAR data
the use of this technique could be complemented with some high-dimensional POLSAR target-decomposition techniques, such as the Freeman Durden decomposition \cite{Freeman_1998_TGRS_963} or the entropy/anisotropy decomposition \cite{Cloude_1997_TGRS_68} or similar.

Nevertheless the proposed models are promising.
Even though initially developed for partial and monostatic POLSAR data,
  it was shown to be applicable to traditional SAR data as well.
Since the models assumptions are quite minimal, they may potentially applicable to bi-static and interferometric data, although that would require further study.

Intuitively, the concept of the determinant being representative of the POLSAR data can be thought of as being similar to the concept of using the magnitude as being representative of a complex number
(in a sense this is also why intensity is widely considered representative of the SAR data complex).
Similar to the determinant, the magnitude observable is also scalar and is not lossless.
To fully represent the complex number, another observable is needed besides the magnitude.
The magnitude however is much more widely used, as it can be shown that its is invariant to a change in the reference frame.
Similarly, the determinant of the POLSAR covariance matrix is also scalar and invariant to a change of polarization basis.
Thus while it is not fully representative of the POLSAR data, it can be said to be highly representative of it.
The important thing is that, it can be used to give a scalar comparison of the multi-dimensional data.
This is also similar to the standard use of the magnitude when two or more complex numbers are to be compared. %, most of the times the magnitude is to be used.

This leads to the application of the models in deriving discrimination measures.
Besides the above properties, the theoretical models may also provide an alternative derivation for the widely used likelihood test statistics in POLSAR.
In view of the models given in Eqns \ref{eqn:determinant_distribution} \& \ref{eqn:log_determinant_distribution},
  the likelihood test statistics exposed in \cite{Conradsen_2003_TGRS_4} and rewritten in Eqns \ref{eqn:ori_likelyhood_stats} \& \ref{eqn:log_likelyhood_stats}
can be simulated as:
\begin{align*}
  \ln{Q} &\sim  k + L_x \Lambda^d_{L_x} + L_y \Lambda^d_{L_y} - (L_x + L_y) \Lambda^d_{(L_x + L_y)} \\
  Q &\sim e^k \frac{(\chi^d_{L_x})^{L_x} \cdot (\chi^d_{L_y})^{L_y}}{(\chi^d_{L_x + L_y})^{L_x + L_y}}   
\end{align*}
where $k = d \left[ (L_x + L_y) \ln(L_x + L_y) - L_x \ln{L_x} - L_y \ln{L_y} \right]$.
As a by-product of this exact derivation, the current paper also proposed several simpler discrimination measures for the common case of $L_x=L_y$.

Similar to the way that other measures of distance can be used to derive POLSAR classifiers \cite{Lee_1999_TGRS}, change detectors \cite{Conradsen_2003_TGRS_4}, edge detectors \cite{Schou_2003_TGRS_20} or other clustering and speckle filtering techniques \cite{Le_2010_ACRS} \cite{Le_2011_ACRS}, 
new detection, classification, clustering or speckle filtering algorithms can be derived using the models presented in this paper.
These applications however are not included in this paper, but rather left as a suggestion for future investigations.
There are a few reasons for this decision.
The first is that the advantages of the proposed models have been demonstrated through a few applications in this paper:
  the first being a faster ENL evaluation for POLSAR data and
  the second being the visual evaluation of POLSAR speckle filters.
The second reason is that such an application would be quite big in scope, and as such would easily distract the reader from the main proposal of this paper.
Last but not least, this paper focuses on the proposal of several scalar and representative statistical models for POLSAR, which results in associated statistical models and discrimination measures.
It does not, technically speaking, focuses on proposing new discrimination measures.
%***IVM remove this single sentence, it detracts from the argument...
%Moreover many of the existing techniques in SAR may be adaptable for POLSAR data, such as visual evaluation of speckle filters.

The advantages of this proposal are thus to be compared against other published models for POLSAR.
Consequently, the methodology to evaluate the result of this proposal is a bit unusual.
While quantitative evaluation is normally employed, in this theoretical model proposal only qualitative evaluation are shown.
This is because compared to existing published observables for POLSAR, the advantage of the proposed observables are qualitative which does not requires numerical quantitative comparisons.
Instead a straightforward binary evaluation would be sufficient.
In comparison to other scalar models for POLSAR, the proposed models for determinant and its log-transformed version are representative are highly representative of the POLSAR data.
This is because these models collapses into the representative SAR intensity when the multi-dimensional data is collapsed into the traditional one-dimensional SAR data.
This is proven in section \ref{sec:sar_special_case_of_polsar}.
No other existing models for POLSAR is known to have this property.

Other advantage is that compared to other proposed scalar models for POLSAR, only these proposed statistical models lead to existing discrimination measures.
This argument is supported by an extensive literature survey in section \ref{sec:lit_review}.
There not only existing published scalar models for POLSAR are reviewed to show that none of them lead to discrimination measure,
  existing discrimination measures for POLSAR are also extensively surveyed leaving virtually no chance of there exists an existing scalar statistical model for POLSAR which leads to existing discrimination measures.
%to comprehensively support the argument. 
%To support these arguments, extensive literature surveys has been carried out for both published scalar models as well as for discrimination measures for POLSAR. Thus given such extensive reviews the above claims are water-tight.

There may have many beneficials implications for this proposal.
One of the main benefits is believed to be that:
  this proposal allows the extension of many of existing SAR data processing technique toward POLSAR.
This is illustrated through an example which extends the SAR speckle filters evaluation technique towards POLSAR.  
%Here a major benefit is illustrated in showing that 
In evaluating SAR speckle filters, the intensity ratio is widely used.
Specifically, in its original multiplicative domain, the ratio of the filtered output and the noisy input image represents the noise being removed.
Assuming a perfect filtering condition, the ratio image should only contain random noise.
Thus a commonly used visual evaluation is to simply plot the ratio image and determine if any image features have also been removed.

@@

To demonstrate this feature, the technique is used to evaluate the performance of 3$\times$3 and 5$\times$5 boxcar POLSAR filters on the AIRSAR Flevoland partial polarimetric data (HH-HV).
  A square 700$\times$700 pixel patch is first extracted from the AIRSAR dataset,
  and the two POLSAR speckle filters are then applied to the patch.
\footnote{While the boxcar filters may not stay in the frontier of POLSAR speckle filtering,
  they are workable here since the focus of this experiment is not on finding the best POLSAR speckle filters available.
  Rather its purpose is to show a quick example of how an existing technique in SAR data processing can be easily adapted for POLSAR data.} 
The log-determinant images of the filtered outputs are displayed in Fig. \ref{fig:visual_eval_part_pol_boxcar_speckle_filters_3x3_vs_5x5}.
At the same time, the residual is computed for both filters, and the images are also displayed in the same figure.

@@

%Ian version: There are many similar SAR data processing techniques such as  the usage of  the mean and variance of the ratio image to quantitatively evaluate speckle filters.
%Evidently these techniques can also be adapted for POLSAR.
%Other examples of such techniques include existing SAR speckle filters, target detectors, land-cover classifiers and edge detectors.

%Hai version: (Ian have not seen this yet)
There are many similar SAR data processing technique that may be adaptable to POLSAR.
Examples in the same topic of evaluating speckle filters include
  the use of the mean and variance of the ratio image, or ENL evaluation to quantitatively evaluate POLSAR speckle filters.
Other topics that may also be connected include: speckle filtering, target detection, surface cover classification, edge detection ... and so on.

Currently while the field of SAR is much more developed than the field of POLSAR,
  the two fields however are quite separated.
Hence many techniques applicable to SAR are not directly applicable to POLSAR.
Under this context, the current proposal provide the bridge between these two separated fields,
  inviting potential extension of various SAR processing techniques to POLSAR data.
%Here a specific example is shown
For instance the earlier example illustrate that while intensity ratio was commonly used to carry out visually evaluation of SAR speckle filters,
  the newly proposed determinant ratio and its log-transformed version (i.e. contrast) are shown capable of visually evaluating POLSAR speckle filters.
We believe there are many more unexplored possibilities for such applications.
%  which is only limited by our imagination.

@@

A number of scalar statistical models for the determinant of the POLSAR covariance matrix are proposed and validated in this paper,
  resulting in several consistent discrimination measures useful POLSAR applications.
The theoretical models are shown to be comprehensive in that
  not only can they provide alternative and simple explanations to a range of theoretical concepts such as POLSAR test statistics or ENL estimation,
  but they can also be considered multi-dimensional extensions of traditional models for one-dimensional SAR data.
The models are also practically useful since they provide several consistent discrimination measures.
Moreover, compared to other scalar statistical models for POLSAR, the proposed models are excellent representations of the multi-dimensional POLSAR data.
Consequently, the derived discrimination measures may be employed in a wide range of applications where a scalar number is required to represent the complex multi-dimensional POLSAR data.
Thus the models are envisioned to form a bridge inviting the adaptation of many existing SAR data processing techniques for POLSAR data.

@@

In this appendix, the following results for SAR intensity $I$ are shown to be special cases of the results given in this paper for the determinant of the POLSAR covariance matrix $det|C_v|$.
Specifically, the following results extend from the authors previous work on single-look SAR \cite{Le_2010_ACRS}, i.e. $d=L=1$, which is considerd a special case. We can state the following:
\begin{eqnarray}
  I &\sim& \bar{I} \cdot pdf \left[ e^{-R} \right] \\
  \log_2{I} &\sim& \log_2{\bar{I}} + pdf \left[ 2^{D-2^D} \right] \\
  \frac{I}{\bar{I}} = \mathbb{R} &\sim& pdf \left[ e^{-R} \right]  \\
  \log_2{I} - \log_2{\bar{I}} = \mathbb{D} &\sim& pdf \left[ 2^De^{-2^D}\ln2 \right]\\
  \log_2{I_1} - \log_2{I_2} = \mathbb{C} &\sim& pdf \left[ \frac{2^c}{(1+2^c)^2} \ln2 \right] \\
  avg(\mathbb{D}) &=& -\gamma / \ln{2} \\
  var(\mathbb{D}) &=& \frac{\pi^2}{6} \frac{1}{ \ln^2{2}} \\
  mse(\mathbb{D}) &=& \frac{1}{\ln^2{2}}( \gamma^2 + \pi^2/6 ) = 4.1161 
\end{eqnarray}
but also the following well-known results are considered for multi-look SAR, i.e. $d=1,L>1$:
  \begin{eqnarray}
I &\sim& pdf \left[ \frac{L^L I^{L-1} e^{-LI/\bar{I}}}{\Gamma(L) \bar{I}^L} \right] \\
N = \ln{I} &\sim& pdf \left[ \frac{L^L}{\Gamma(L)} e^{L(N-\bar{N})-Le^{N-\bar{N}}} \right]
  \end{eqnarray}
It will be shown that all of these results are special cases of the result derived previously and rewritten below:
\begin{eqnarray}
  |C_v| &\sim& \frac{|\Sigma_v|}{(2L)^d} \prod^{d-1}_{i=0} \chi^2(2L-2i)  \label{eqn:polsar_det_cov_dist} \\
  \ln{|C_v|} &\sim& \ln{|\Sigma_v|} + \sum^{d-1}_{i=0} \Lambda(2L-2i) - d \cdot \ln{2L} \label{eqn:polsar_log_det_cov_dist} 
\end{eqnarray}
\begin{eqnarray}
  \frac{|C_v|}{|\Sigma_v|} = \mathbb{R} &\sim& \frac{1}{(2L)^d} \prod^{d-1}_{i=0} \chi^2(2L-2i) \label{eqn:polsar_ratio_det_cov_dist} \\
  \ln{|C_v|} - \ln{|\Sigma_v|} = \mathbb{D} &\sim& \sum^{d-1}_{i=0} \Lambda(2L-2i) - d \cdot \ln{2L} \label{eqn:polsar_dispersion_log_det_cov_dist} \\ 
  \ln{|C_{1v}|} - \ln{|C_{2v}|} = \mathbb{C} &\sim& \sum^{d-1}_{i=0} \Delta(2L-2i)
\end{eqnarray}
\begin{eqnarray}
  avg(\mathbb{D}) &=& \sum^{d-1}_{i=0} \psi^0(L-i) - d \cdot \ln{L} \label{eqn:polsar_dispersion_averages} \\
  var(\mathbb{D}) &=& \sum^{d-1}_{i=0} \psi^1(L-i) \label{eqn:polsar_dispersion_variance} \\
  mse(\mathbb{D}) &=& \left[ \sum^{d-1}_{i=0} \psi^0(L-i) - d \cdot \ln{L} \right]^2 +  \sum^{d-1}_{i=0} \psi^1(L-i) \label{eqn:polsar_dispersion_mse}
\end{eqnarray}

@@

The output of the first phase, in the general case, is applicable to multi-look SAR data, where $d=1$ but $L>1$.
The second phase simplifies the multi-look results into single-look results, which will match those presented in our previous work \cite{Le_2010_ACRS}.
Mathematically, it means setting $L=1$ in the multi-look result
  and converting from the natural logarithmic domain used in this paper to the base-2 logarithm used in \cite{Le_2010_ACRS} (base-2 was chosen in the previous paper to simplify the computation).

@@

Now setting $L=1$, these results become:
\begin{eqnarray}
    I &\sim& pdf \left[ \frac{ e^{x/\bar{I}}}{ \bar{I}} \right] \\
    \frac{I}{\bar{I}} = \mathbb{R} &\sim& pdf \left[ e^{-x} \right] 
\end{eqnarray}
which is the same as stated in \cite{Le_2010_ACRS}, demonstrating that the previous work is a special case of the more generic POLSAR forms.

@@

Setting $L=1$ for Single-Look SAR we have
\begin{equation}
  \mathbb{D} \sim pdf \left[ e^{x-\gamma-e^{x-\gamma}} \right]
\end{equation}
due to: $\psi^0(1)=-\gamma$ and $\Gamma(1)=1$
with $\gamma$ being the Euler Mascheroni constant (which equals $0.5772$). 
In base-2 logarithm domain, invoking the variable change theorem:
\begin{eqnarray*}
  \mathbb{D}_2 &=& \log_2{I} - avg(\log_2{I}) = \mathbb{D}/\ln{2} \\
  \mathbb{D}_2 &\sim& pdf \left[ e^{x-\gamma-e^{x-\gamma}} \cdot \frac{dx}{dt} \right]_{x=t \cdot \ln2}
\end{eqnarray*}
Thus we have
\begin{equation}
  \mathbb{D}_2 \sim pdf \left[ e^{-(2^xe^{-\gamma})} (2^xe^{-\gamma}) \ln2 \right]
\end{equation}
which is consistent with the results found in our previous work \cite{Le_2010_ACRS}.
